Interacting with Modal Logics \\ in the Coq Proof Assistant

### 2015-01-15 Papers/InProgress/ModalLogicCoq/ModalCoq.tex
Improve and compress language
Conflicts:
	Papers/InProgress/ModalLogicCoq/ModalCoq.pdf

However, one may wonder whether the embedding approach is adequate
also for \emph{interactive} reasoning, when the user proves theorems
by interacting with a proof assistant such as \Coq\footnote{The \Coq
proof assistant was chosen merely because of the authors' greater
familiarity with the tactic language of this system. Nevertheless, the
techniques presented and discussed here are likely to be useful for
other proof assistants, such as \texttt{Isabelle} \cite{Isabelle} or
\texttt{HOL Light} \cite{HOLLight}.}. The main goal and novelty
of this paper is to study this question. Our answer is positive.

One major initial concern was whether the embedding could be a
disturbance to the user. Fortunately, by using \Coq's \texttt{Ltac}
tactic language, we were able to define intuitive new tactics that
hide the technical details of the embedding from the user. The
resulting infra-structure for modal reasoning within \Coq (as
described in Section \ref{sec:Embedding}) provides a user experience where
modalities can be handled transparently and straightforwardly.
Therefore, a user with basic knowledge of modal logics and \Coq's
tactics should be able to use (and extend) our implementation with no
excessive overhead.

@@

The simpler ontological argument of Anselm has been automatically
verified with \texttt{PVS} by Rushby \cite{Rushby} and with 
first-order theorem provers by Oppenheimer and Zalta \cite{Zalta}. G\"odel's
argument was automatically verified in our previous work on embedding-
based fully automated modal theorem proving \cite{arXiv,AFP}. This
paper presents the first fully interactive and detailed formalization
of this proof in a proof assistant. Hence we here reveal the proof
structure on a cognitively adequate level of
detail. Such detailed proof structure has been hidden in our
other papers on the subject due to the use of automated theorem
provers. In comparison to \cite{Rushby,Zalta}, however, our
contribution stands out with its surprising technical 
simplicity and elegance.

In addition to philosophy, propositional and quantified modal logics 
have (potential) applications in various other fields as well, including, for instance, 
verification, artificial intelligence
agent technologies, law and linguistics (cf.~\cite{ModalLogic} and the references therein).
Therefore, the main contribution described in this paper -- convenient techniques for leveraging a powerful proof assistant such as \Coq for interactive reasoning for modal logics -- may serve as a starting point for
many interesting projects.

>>>>

However, one may wonder whether the embedding approach is adequate
also for \emph{interactive} reasoning, when the user proves theorems
by interacting with a proof assistant such as \Coq\footnote{The \Coq
proof assistant was chosen because of the authors' greater
familiarity with the tactic language of this system. Nevertheless, the
techniques presented here are likely to be useful for
other proof assistants (e.g \texttt{Isabelle} \cite{Isabelle},
\texttt{HOL-Light} \cite{HOLLight}).}. The main goal and novelty
of this paper is to study this question. Our answer is positive.

One major initial concern was whether the embedding could be a
disturbance to the user. Fortunately, by using \Coq's \texttt{Ltac}
tactic language, we were able to define intuitive new tactics that
hide the technical details of the embedding from the user. The
resulting infra-structure for modal reasoning within \Coq (as
described in Sections \ref{sec:Embedding} and \ref{sec:Tactics}) provides a user experience where
modalities can be handled transparently and straightforwardly.
Therefore, a user with basic knowledge of modal logics and \Coq's
tactics should be able to use (and extend) our implementation with no
excessive overhead.

@@

The ontological argument of Anselm has been automatically
verified with \texttt{PVS} by Rushby \cite{Rushby} and with 
first-order theorem provers by Oppenheimer and Zalta \cite{Zalta}. 
In comparison, our
contribution stands out with its surprising technical 
simplicity and elegance, despite the greater complexity
of G\"odel's argument.

G\"odel's argument was automatically verified in our previous work on fully automated modal theorem proving based on embedding \cite{arXiv,AFP}. 
This paper presents the first fully interactive and detailed formalization
of this proof in a proof assistant. The proof
structure, which has been hidden in our
other papers on the subject due to the use of automated theorem
provers, is revealed here on a cognitively adequate level of
detail.

In addition to philosophy, propositional and quantified modal logics 
have (potential) applications in various other fields, including, for instance, 
verification, artificial intelligence
agent technologies, law and linguistics (cf.~\cite{ModalLogic} and the references therein).
Therefore, the main contribution described in this paper -- convenient techniques for leveraging a powerful proof assistant such as \Coq for interactive reasoning for modal logics -- may serve as a starting point for
many interesting projects.
